---
title: 'Advanced Time Series Analysis: Computer Exercise 2'
author: "Anders Launer Bæk (s160159)"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
rm(list = ls())

knitr::opts_chunk$set(echo = FALSE, 
                      include = TRUE,
                      warning = FALSE,
                      fig.width = 8, fig.height = 4,
                      fig.show='hold', fig.align='center',
                      
                      eval = TRUE, 
                      tidy = TRUE, 
                      dev = 'pdf', 
                      cache = TRUE, fig.pos = "th!")

kable_format <- list(small.mark = ",",
                     big.mark = ',',
                     decimal.mark = '.',
                     nsmall = 3,
                     digits = 3,
                     scientific = FALSE,
                     big.interval = 3L)

library(ggplot2)
theme_TS <- function(base_size = 9, base_family = "", face = "plain"){
  theme_bw(base_size = base_size, base_family = base_family) %+replace%
    theme(panel.background = element_blank(), 
          panel.border = element_blank(),
          panel.grid = element_blank(),
          axis.text = element_text(size = base_size, face = face, family = base_family),
          axis.title = element_text(size = base_size, face = face, family = base_family),
          legend.text = element_text(size = base_size, face = face, family = base_family))
}

fig <- local({
    i <- 0
    ref <- list()
    list(
        cap=function(refName, text) {
            i <<- i + 1
            ref[[refName]] <<- i
            #paste("Figure ", i, ": ", text, sep="")
            paste(text, sep="")
        },
        ref=function(refName) {
            ref[[refName]]
        })
})
# tab <- local({
#     i <- 0
#     ref <- list()
#     list(
#         cap=function(refName, text) {
#             i <<- i + 1
#             ref[[refName]] <<- i
#             paste("Table ", i, ": ", text, sep="")
#         },
#         ref=function(refName) {
#             ref[[refName]]
#         })
# })

library(rgl)
```
## Part 1
There is generated $n=1000$ noise samples from a $x \sim \mathcal{N}(0,\,1)$ which is used as the noise input for `t` in all simulations in part one. 

```{r, fig.cap="\\label{fig_ex_1}Plot of sampled noise."}
## Number of samplepoints
n <- 1000
## Uniform distributed x
#x <- runif(n,-1,1)
## Errors
# r <- rnorm(n, mean = 0, sd = 1)
# r <- rnorm(3000, mean = 0, sd = 1)
#save(r, file = "./noise.Rda")
load(file = "./noise.Rda")
#data <- data.frame(t = 1:n, x = x, noise = r)
data <- data.frame(t = 1:n, noise = r[1:n])

# ggplot(data) +
#    geom_point(aes(x = t, y = noise, color = "X ~ N(0, 1)"), alpha = 1/2) +
#    labs(x = "t (time)", y = "y(t)", color = "") + 
#    theme_TS()
```

The equation below, eq. \ref{eq_1_2a} and eq. \ref{eq_1_2b}, determines the parameter set for the SETAR(2,1,1) model (eq. \ref{eq_1_SETAR_r}).

\begin{equation}
a_0 = [2.0, -1.0]
\label{eq_1_2a}
\end{equation}
\begin{equation}
a_1 = [0.6, -0.9]
\label{eq_1_1b}
\end{equation}

### Simulation of the SETAR(2,1,1)

The Self-Exciting Threshold AR (SETAR) model is given by eq. \ref{eq_1_SETAR}.

\begin{equation}
X_t = a_0^{(J_t)} + \sum_{i = 1}^{k_{(J_t)}} a_i^{(J_t)} X_{t-i}+\epsilon^{(J_t)}
\label{eq_1_SETAR}
\end{equation}
where $J_t$ are regime processes. The complete model are defined in eq. \ref{eq_1_SETAR_r}.

\begin{equation}
X_t = \left\{ \begin{matrix} a_{0,1} + a_{1,1} X_{t- 1} + \epsilon_t & for & X_{t-1} \leq0 \\  a_{0,2} + a_{1,2} X_{t- 1} + \epsilon_t &  for & X_{t-1} >0 \end{matrix} \right\}
\label{eq_1_SETAR_r}
\end{equation}

The model $X_t$ (eq. \ref{eq_1_SETAR_r}) has been simulated with two different set of parameters (eq. \ref{eq_1_2a} - eq. \ref{eq_12_2b}) and its simulations are plotted in fig. \ref{fig_ex_2}. 


```{r, fig.cap="\\label{fig_ex_2}Two simulated SETAR(2,1,1) models using $par_1$ and $par_2$."}

# define parameter
a0 <- c(2.0, -1.0) # off set
a1 <- c(0.6, -0.9) # slope


## Make a time series y with a regime model
data$y_SETAR <- rep(NA,n)
data$y_SETAR[1] <- data$noise[1]

for(t in 2:n) {
  if(data$y_SETAR[t - 1] <= 0) {
    data$y_SETAR[t] <- a0[1] + a1[1] * data$y_SETAR[t-1] + data$noise[t]
  } else {
    data$y_SETAR[t] <- a0[2] + a1[2] * data$y_SETAR[t-1] + data$noise[t]
  }
}

#
data$y_SETAR_1 <- rep(NA, n)
data$y_SETAR_1[2:n] <- data$y_SETAR[1:(n-1)]

# plot data
ggplot(data) +
  geom_point(aes(x = y_SETAR_1, y = y_SETAR, color = "SETAR(2,1,1)"), alpha = 1/2) +
  labs(x = "y(t - 1)", y = "y(t)", color = "") +
  theme_TS()
```

Fig. \ref{fig_ex_2} shows the plot of the SETAR(2,1,1) model with the two different parameter sets. 
\begin{itemize}
\item For both model it is possible to differentiate between the regimes and their transitions.
\item It is also possible to see the inverse properties of the slop for the two models.
\item Both models are using different offsets where the transition are most separated in the model which is using $par_2$.
\end{itemize}


### Estimate the parameters using conditional least squares

see section 5.5 on page 115

```{r, echo=TRUE}
RSSSetar <- function(theta) {
  p1 <- theta[1]
  p2 <- theta[2]
  ## Calculate the objective function value
  obj <- 1
  
  #
  return(obj)
}
```

```{r, fig.cap="\\label{fig_ex_3}Plot of sampled noise."}






```

